\textbf{Urban mobility and socioeconomic deprivation in Latin America after COVID-19}

### 2024-10-23 manuscript/manuscript/manuscript.qmd
manuscript-section-results-i

Drawing on a dataset of 213 million observations from Meta-Facebook users' mobile location data, we aim to assess socioeconomic differences in the extent and persistence of decline in urban mobility in Argentina, Chile, Colombia and Mexico during and after the COVID-19 pandemic from March 2020 to March 2023.
We use Meta-Facebook data to measure origin-destination flows from March 2020 to May 2022, and Meta Prophet time-series forecasting machine learning algorithm [@prophet17] to predict origin-destination flows from June 2022 to March 2023.
<!--# Perhaps say how predictions have been validated --> We use Functional Urban Areas (FUAs) boundaries from the Global Human Settlement Layer (GHSL), developed by the European Comission's Joint Research Centre [@GHSL19] to define urban areas; and the Global Gridded Relative Deprivation Index (GRDI) developed by NASA's Socioeconomic Data and Applications Centre [@RDI2022] from sociodemographic and satellite data inputs.
Building on existing evidence [e.g. @roweCalafiore2022; @wang2022], we hypothesised that (1) urban mobility has recovered returning to the pre-pandemic baseline level of movement as nonpharmaceuthical restrictions were lifted; and, that (2) socioeconomic differences in urban mobility have endured the pandemic reflecting deep societal inequalities as knowledge-intensive businesses adopt hybrid working.
<!--# Once we have a more definitive version of the results, highlight our novel contributions here. -->\

@@

@@

<!--![Administrative units within functional urban areas in Argentina, Chile, Colombia and Mexico, coloured according to the average Relative Deprivation Index.](figures/maps.pdf)!-->

## The heterogeneous impact of COVID-19 on urban mobility

We analyse the evolution of the percentage change in the number of movements with respect to a baseline period prior to the pandemic.
Specifically, we focus the analysis on short-distance movements in urban areas to represent local and routine mobility [@owen1992migration], so only movements covering a distance of at most 70 km are considered.
For a movement to be classified as urban it needs to start or end within a functional urban area from Argentina, Chile, Colombia and Mexico.
The observed data is available for a two-year period starting in April 2020, just after the first wave of COVID-19 pandemic cases, and ending in March 2022.
After 2022 no observations are available, however, we generate a 12-month forecast up to March 2023 in order to gain a better understanding of the recovery trends.

@@

## The most deprived areas experienced the smallest drop in mobility relative to pre-pandemic levels

@@

## WorlPpop population data

@@

In our analysis, we use the Global Gridded Relative Deprivation Index (GRDI), Version 1 (GRDIv1) data set as a measure of socioeocnomic deprivation.
The GRDI data is made available via NASA's Socioeconomic Data and Applications Centre (SEDAC), at a spatial resolution of 30 arc-seconds, or 1 km$^2$ approximately.
The index quantifies the relative levels of multidimensional deprivation and poverty, where a value of 100 represents the highest level of deprivation and a value of 0 the lowest.
We perform a spatial join of the Facebook spatial units and the gridded relative deprivation data and compute the average RDI corresponding to each of the Facebook spatial units.

>>>>

Drawing on a dataset of 213 million observations from Meta-Facebook users' mobile location data, we aim to assess socioeconomic differences in the extent and persistence of decline in urban mobility in Argentina, Chile, Colombia and Mexico during and after the COVID-19 pandemic from March 2020 to March 2023.
We use Meta-Facebook data to measure origin-destination flows from March 2020 to May 2022, and Meta Prophet time-series forecasting machine learning algorithm [@prophet17] to predict origin-destination flows from June 2022 to March 2023.
<!--# Perhaps say how predictions have been validated --> We use Functional Urban Areas (FUAs) boundaries from the Global Human Settlement Layer (GHSL), developed by the European Comission's Joint Research Centre [@GHSL19] to define urban areas; and the Global Gridded Relative Deprivation Index (GRDI) developed by NASA's Socioeconomic Data and Applications Centre [@RDI2022] from sociodemographic and satellite data inputs.
Building on existing evidence [e.g. @roweCalafiore2022; @wang2022], we hypothesised that (1) urban mobility has recovered returning to the pre-pandemic baseline level of movement as nonpharmaceuthical restrictions were lifted; and, that (2) socioeconomic differences in urban mobility have endured the pandemic reflecting deep societal inequalities as knowledge-intensive businesses adopt hybrid working.
<!--# Once we have a more definitive version of the results, highlight our novel contributions here. -->

@@

## Analysing the variability in the impact of COVID-19 on mobility across population groups

The focus of this work is on capturing how mobility patterns for different population groups have been impacted by COVID-19 in Latin America.
For this purpose, we consider populations groups by classifying their local areas into categories according to their population density and their socioeconomic level.
These two criteria for classification are chosen due to their relevance for ...

We use five population density categories and three socioeconomic categories defined according to the relative deprivation index.
These categories are used to classify the spatial units of analysis, following the methodology defined in section X.
The geographic distribution of categories for each of the countries is displayed in Fig X.a), Fig X.a) and Fig X.a) for Argentina, Chile and Colombia respectively.
Fig X.b), Fig X.b) and Fig.b) show, for each country, the population distribution into the categories of analysis, considering, on the one hand, Worldpop population estimations and on the other hand, Facebook population, based on user counts during the pre-pandemic baseline period (see section X of the methodology for further details).
For the three countries we observe discrepancies in the population distribution according to WorldPop and Facebook data, which reflect the extent to which the Facebook data tends to overrepresent certain population groups with respect to the Worldpop data whilst underrepresenting others.
For example, for Argentina and Chile, the proportion of population living in areas of low population density is higher according to Worldpop data, hence indicating that Facebook data underrepresents populations in these regions.
Similarly, in those countries, Facebook data shows and overrepresentation of the most affluent socioeconomic group.

Biases in Facebook data arise for a variety of reasons.
EXPLAIN.
Finally, Meta shares Facebook data for active user population and flows that has been spatially and temporally aggregated.
Their data aggregation methodology is designed so that records corresponding to low counts are not included in the dataset in order to preserve the privacy of users.
Analysing aggregated Facebook data in its raw format can therefore result in an underrepresentation of groups that produce low data counts.
To counter these source of bias in the Facebook data, we propose a methodology to process the data, which is fully described in the section X.

The analysis of mobility is done using Facebook data that has been pre-processed according to the proposed methodology for bias correction.
Fig X.c), Fig X.c) and X.c) show, for Argentina, Chile and Colombia respectively, changes in the intensity of movement relative to pre-pandemic baseline levels, at two points during the period of analysis, May 2020 and March 2022.
The patterns are consistent across the three countries in both dimensions of I'M HERE!!!!!

## COVID-19 unevenly impacted human mobility across population categories in Latin America

@@

We analyse the evolution of the percentage change in the number of movements with respect to a baseline period prior to the pandemic.
Specifically, we focus the analysis on short-distance movements in urban areas to represent local and routine mobility [@owen1992migration], so only movements covering a distance of at most 70 km are considered.
For a movement to be classified as urban it needs to start or end within a functional urban area from Argentina, Chile and Colombia.
The observed data is available for a two-year period starting in April 2020, just after the first wave of COVID-19 pandemic cases, and ending in March 2022.
After 2022 no observations are available, however, we generate a 12-month forecast up to March 2023 in order to gain a better understanding of the recovery trends.

@@

## Recovery trajectories towards baseline mobility levels are heterogeneous across population categories in Latin America

@@

## WorldPpop population data

@@

We use the Global Gridded Relative Deprivation Index (GRDI), Version 1 (GRDIv1) data set as a measure of socioeocnomic deprivation.
The GRDI data is made available via NASA's Socioeconomic Data and Applications Centre (SEDAC), at a spatial resolution of 30 arc-seconds, or 1 km$^2$ approximately.
The index quantifies the relative levels of multidimensional deprivation and poverty, where a value of 100 represents the highest level of deprivation and a value of 0 the lowest.
We perform a spatial join of the Facebook spatial units and the gridded relative deprivation data and compute the average RDI corresponding to each of the Facebook spatial units.

### 2024-10-23 manuscript/manuscript/manuscript.tex
manuscript-section-results-i

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2.54cm,right=2.54cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{lipsum} \usepackage{libertine}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\date{}
\begin{document}
\maketitle
\begin{abstract}
COVID-19 triggered a reduction in the number and extent of peoples
movements. Existing evidence suggests that while the impact of the
pandemic on mobility was widespread, the extent of this impact was
unequally felt across socioeconomic groups in the early stages of the
pandemic. Here, we find that in the long term, the most deprived
locations have experienced a more accelerated recovery towards
pre-pandemic levels of mobility. Furthermore, the socioeconomic
disparities in the patterns of mobility triggered by the first outbreak
of COVID-19, have persisted as of April 2023. These findings are based
on the analysis of time-series mobility data corresponding to X urban
areas from Latin American countries collected from Meta-Facebook users
upon their consent. Our research highlights the importance of timely
mobility data with high spatiotemporal resolution to understand the
long-term effects of the pandemic and to inform equitable policy
responses that address societal challenges in urban areas.
\end{abstract}

@@

\begin{figure}[H]

{\centering \includegraphics{figures/maps.pdf}

}

\caption{Administrative units within functional urban areas in
Argentina, Chile, Colombia and Mexico, coloured according to the average
Relative Deprivation Index.}

\end{figure}%

@@

\begin{figure}[H]

{\centering \includegraphics{figures/prediction-rdi-band.pdf}

}

\caption{Patterns of recovery for urban mobility in administrative units
from Argentina, Chile, Colombia and Mexico.}

\end{figure}%

@@

\begin{figure}[H]

{\centering \includegraphics{figures/regression-evo-nobackground.pdf}

}

\caption{Evolution of the parameters characterising the relationship
between relative deprivation index and the percentage change in the
number of movements relative to the pre-pandemic baseline level of
mobility.}

\end{figure}%

@@

\section{Methods}\label{sec-methods}

\subsection{Meta-Facebook movement
data}\label{meta-facebook-movement-data}

@@

\subsection{Spatiotemporal data
aggregation}\label{spatiotemporal-data-aggregation}

Since the focus of this work is on urban mobility, we focus the analysis
on Functional Urban Areas (FUAs), defined by the Global Human Settlement
Layer (GHSL). The spatial extent of the FUAs is often large and may
include several towns and neighbourhoods displaying a variety of
socioeconomic characteristics, hence using FUAs as the spatial units of
aggregation would considerably mask the heterogeneity in the mobility
patterns. While the original unit of aggregation for the movement data,
i.e.~the tiles from the Bing Maps Tile System, offers the highest degree
of spatial granularity available, the interpretation of findings from an
analysis based on administrative units is often more valuable. For this
reason, we perform a spatial join to aggregate the movement data into
administrative units at the GADM level 2 or 3. Only flows of people
starting or ending within the boundaries of a FUA are considered.

While the original movement data is originally aggregated into 8-hour
windows, this resolution is too fine for our analysis. Since the
analysis is focused on the longer-term evolution of patterns of urban
mobility, we aggregate the movement data by month.

In our analysis, we use the Relative Deprivation Index (RDI) as a
measure of socioeocnomic deprivation. The RDI data is made available via
NASA's Socioeconomic Data and Applications Centre (SEDAC), with a
spatial resolution of 1km pixels. We perform a spatial join of the
gridded data and the administrative units and compute the average RDI
within each of these administrative units.

\subsection{Time series analysis}\label{time-series-analysis}

For the countries included in the analysis, the movement data is
available until March 2022. In order to gain a better understanding of
the recovery trends after the pandemic, we generate a 12-month forecast
up to March 2023. This is done using Prophet\textsuperscript{27}, a
procedure developed by Meta's Core Data Science team to forecast time
series data based on an additive model. Prophet stands out from other
forecasting methods due to its ability to capture recurring seasonal
trends inherent in mobility data, such as fluctuations during holidays,
or changes in the general trend due to the impact of particular
circumstances, such as the COVID-19 pandemic. It is particularly
intuitive to use compared to other models due to its automated
trend-detection capabilities, making it accessible to users with no
expertise on the underlying model, while being robust to missing data
and outliers. For the analysis in this paper, yearly seasonality effects
are included as they yield the most realistic predictions.

Figures 2 and 3 have been created by removing statistical outliers. We
define these as the administrative units which, at any given month, have
a z-score greater than 4 for the percentage change in the number of
movements. Furthermore, a Savitzky-Golay filter is applied to smooth the
time series data, using a length of 4 units for the filter window and
order 2 polynomials to fit the samples.

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2.54cm,right=2.54cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{lipsum} \usepackage{libertine}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\date{}
\begin{document}
\maketitle
\begin{abstract}
COVID-19 triggered a reduction in the frequency and extent of people's
movement. Existing evidence suggests that while the impact of the
pandemic on mobility was widespread, the extent of this impact was
unequally felt across socioeconomic groups in the early stages of the
pandemic. Here, we find that the most deprived locations have
experienced a more accelerated recovery towards pre-pandemic levels of
mobility in the long term. Furthermore, the socioeconomic disparities in
the patterns of mobility triggered by the first outbreak of COVID-19,
have persisted as of April 2023. These findings are based on the
analysis of time-series mobility data corresponding to X urban areas
from Latin American countries collected from Meta-Facebook users upon
their consent. Our research highlights the importance of timely mobility
data with high spatiotemporal resolution to understand the long-term
effects of the pandemic and to inform equitable policy responses that
address societal challenges in urban areas.
\end{abstract}

@@

@@

\textless!- --\href{figures/prediction-rdi-band.pdf}{Patterns of
recovery for urban mobility in administrative units from Argentina,
Chile, Colombia and Mexico.} ---!\textgreater{}

@@

@@

\section{Data and Methods}\label{sec-methods}

\subsection{Meta-Facebook data}\label{meta-facebook-data}

@@

\subsection{WorlPpop population data}\label{worlppop-population-data}

We used data from WorldPop\textsuperscript{47} to classify the spatial
units of analysis according to their level of urbanisation, and to
estimate missing baseline values in the Facebook population data.
WorldPop offers open access gridded population estimates at a resolution
as small as 3 arc-seconds approximately 100m and 1km at the Equator,
respectively. WorldPop produces these gridded datasets using top-down
(i.e.~dissagregating administrative area counts into smaller grid cells)
or bottom-up (i.e.~interpolating data from counts from sample locations
into grid cells) approaches. For the purposes of this work, we use
gridded population data at a resolution of 1km\(^2\) in raster format.
We perform a spatial join of the Facebook spatial units (Bing tiles)
with the gridded population data and compute the sum of Worldpop
populations corresponding to each of the Facebook spatial units.

\subsection{Socioeconomic deprivation
data}\label{socioeconomic-deprivation-data}

In our analysis, we use the Global Gridded Relative Deprivation Index
(GRDI), Version 1 (GRDIv1) data set as a measure of socioeocnomic
deprivation. The GRDI data is made available via NASA's Socioeconomic
Data and Applications Centre (SEDAC), at a spatial resolution of 30
arc-seconds, or 1 km\(^2\) approximately. The index quantifies the
relative levels of multidimensional deprivation and poverty, where a
value of 100 represents the highest level of deprivation and a value of
0 the lowest. We perform a spatial join of the Facebook spatial units
and the gridded relative deprivation data and compute the average RDI
corresponding to each of the Facebook spatial units.

\subsection{Classification of tiles according to level of urbanisation
and socioeconomic
deprivation}\label{classification-of-tiles-according-to-level-of-urbanisation-and-socioeconomic-deprivation}

\subsection{Processing Facebook data}\label{processing-facebook-data}

To ensure the privacy and anonymity of the users' data, Meta removes
information corresponding to data entries where the population or
movement count is less than 10 for a specific time or day, retaining
only information about the percentage change in the number of counts
with respect to the baseline period is retained\textsuperscript{46}.
Consequently, we observe many tiles where the population or movement
count for either the baseline or crisis period are blank.

\subsubsection{Facebook population data}\label{facebook-population-data}

To imput Facebook population baseline values, we first identify all the
baseline values that are available for each spatial unit and weekday (of
the three available time windows, we only consider one per day). We then
estimate the missing baseline values for each weekday, based on a linear
model for the relationship between the Worldpop population and the
Facebook population counts, which are fitted using ordinary least
squares. This is illustrated in Supplementary Figure
\ref{fig:imput-baseline-population}.

We then use the complete baseline Facebook population values to compute
missing Facebook population counts during the crisis period. This is
possible because, as mentioned above, Meta reports the percentage change
in the number of counts with respect to the baseline, even if the counts
are not reported due to low value.

\subsubsection{Facebook movement data}\label{facebook-movement-data}

The imputation of Facebook movement baseline values is done according to
a spatial interaction model (see e.g.\textsuperscript{48}). We
considered the population flow between an origin and a destination tile,
and model this variable as a function of the Facebook population count
at the origin tile on the same weekday, the Facebook population count at
the destination on the same weekday and the distance between origin and
destination; we also included indicator variables to capture the day of
the week and the pair of population density classes corresponding to the
origin and destination tiles. Mathematically, this model can be
expressed as \begin{equation}
\mu_{ijw} = \beta_0 + \beta_1pop_{iw} +\beta_2pop_{jw} + \beta_3d_{ij} + \beta_4W + \beta_5C + \varepsilon
\end{equation}

where \(\mu_{ijw}\) is the expectation of the flow of people from tile
\(i\) to tile \(j\) on the weekday \(w\); \(\beta_0\) is an intercept
\(pop_{iw}\) and \(pop_{jw}\) are the Facebook population counts at the
origin and destination on weekday \(w\) during the baseline period,
\(d_{ij}\) is the distance between origin and destination, \(W\) is a
series of indicator variables capturing the day of the week and
similarly \(C\) is a series of indicator variables reflecting the pair
of population density classes for the origin and destination tiles,
resulting in X pairs (10 classes × 10 classes minus one so it is not
collinear with \(\beta_0\)), \(\beta_{0,1,2,3,4,5}\) are model
parameters to be estimated from the observed data. The error term is
denoted by \varepsilon.

To estimate the model parameters, we used a count data regression model.
Specifically, we used a negative binomial regression which is a
generalised linear model (GLM) where overdispersion of the error term is
assumed, i.e.~the variance exceeds the mean.

We compute missing Facebook movement counts during the crisis period by
considering the complete Facebook movements baseline and the percentage
change in the number of counts with respect to the baseline, which is
reported in the Facebook Movements datasets even when the count is not
reported due to its low value.

\subsubsection{Correction factors??? For
representativeness}\label{correction-factors-for-representativeness}

\subsection{Data analysis}\label{data-analysis}

@@

\bibitem[\citeproctext]{ref-tatem2017}
\CSLLeftMargin{47. }%
\CSLRightInline{Tatem, A. J.
\href{https://doi.org/10.1038/sdata.2017.4}{WorldPop, open data for
spatial demography}. \emph{Scientific Data} \textbf{4}, (2017).}

\bibitem[\citeproctext]{ref-rowe2022a}
\CSLLeftMargin{48. }%
\CSLRightInline{Rowe, F., Lovelace, R. \& Dennett, A.
\href{http://dx.doi.org/10.31219/osf.io/xcdms}{Spatial interaction
modelling: A manifesto}. (2022).}
